---
title: "In-class Exercise 07"
author: "Ng Meng Ye"
---

# Geographically Weighted Regression

### 

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary,ggstatsplot)
```

URA Master Plan 2014’s planning subzone boundaries

```{r}
mpsz = st_read(dsn = "C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex07/data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz_svy21)
```

```{r}
st_bbox(mpsz_svy21) #view extent
```

```{r}
condo_resale = read_csv("C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex07/data/aspatial/Condo_resale_2015.csv")
```

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

North side of equatorial

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

```{r}
summary(condo_resale)

```

need to assign

```{r}
condo_resale_sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

```{r}
head(condo_resale_sf)
```
In hands-on exercise,we use corrplot to check multicollinearity, but it runs correlation first

Here we use ggstatplot, only one line of code

```{r}
ggcorrmat(condo_resale[,5:23])
```
perform VIF to check multicollinearity

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD+LEASEHOLD_99YR, 
                data=condo_resale_sf)
summary(condo.mlr)
```
lm = object class in R
fitted values = estimated selling price


use olsrr to perform regression model, can use to perform stepwise regression, directly consume lm object

```{r}
ols_regress(condo.mlr)
```

should perform this to exclude those not significant variable

```{r}
ols_vif_tol(condo.mlr)
```
only few are greater than 0.5

forward stepwise regression,
put in 1 after 1


```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo.mlr,
  p_val = 0.05,
  details = FALSE
)
```

```{r}
plot(condo_fw_mlr)
```
```{r}
ggcoefstats(condo.mlr,
            sort = "ascending")
```
backward stepwise regression
put in everything first, move out one after another

```{r}
ols_plot_resid_fit(condo_fw_mlr$model)
```
hover around the red line, conform the assumption, most

```{r}
ols_plot_resid_hist(condo_fw_mlr$model)
```

```{r}
ols_test_normality(condo_fw_mlr$model)
```
### Test for Spatial Autocorrelation

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>% 
  rename('FW_MLR_RES'='condo_fw_mlr$model$residuals')
```

```{r}
condo_resale_sf <- cbind(condo_resale_sf,
                         mlr_output$FW_MLR_RES) %>% 
  rename('MLR_RES' = 'mlr_output.FW_MLR_RES')
```

```{r}
tmap_mode('view')
tm_shape(mpsz)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) + 
tm_shape(condo_resale_sf) +
  tm_dots(col = "MLR_RES",
          alpha = 0.5,
          style = "quantile")
tmap_mode("plot")
```

### Spatial stationary test

```{r}
library(sfdep)
condo_resale_sf <- condo_resale_sf %>% 
  mutate(nb=st_knn(geometry, k=6,
                   longlat=FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale_sf$MLR_RES,
                  condo_resale_sf$nb,
                  condo_resale_sf$wt,
                  alternative = "two.sided",
                  nsim=99)
```
### Building Fixed Bandwidth GWR model

```{r}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale_sf, 
                          kernel = 'gaussian', 
                          adaptive= FALSE, 
                          longlat = FALSE)
```
```{r}
gwr_fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + 
                         AGE    + PROX_CBD + PROX_CHILDCARE + 
                         PROX_ELDERLYCARE   +PROX_URA_GROWTH_AREA + 
                         PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH +
                         PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                         NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                       data=condo_resale_sf, 
                       bw=bw.fixed, 
                       kernel = 'gaussian', 
                       longlat = FALSE)
```

```{r}
gwr_fixed
```
The report shows that the AICc of the gwr is 42263.61 which is significantly smaller than the global multiple linear regression model of 42967.1.



### Building Adaptive Bandwidth GWR model
```{r}
bw_adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale_sf, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```
30 is the recommended data points to be used


```{r}
gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale_sf, bw=bw.fixed, 
                          kernel = 'gaussian', 
                          longlat = FALSE)
```

```{r}
gwr.adaptive
```
Converting SDF into sf data.frame

```{r}
gwr_adaptive_output <- as.data.frame(
  gwr.adaptive$SDF) %>% 
  select(-c(2:15)
)
```

```{r}
gwr_sf_adaptive <- cbind(condo_resale_sf,
                         gwr_adaptive_output)
```

```{r}
glimpse(gwr_sf_adaptive)
```
```{r}
summary(gwr.adaptive$SDF$yhat)
```
### Visualize local R2

```{r}
tmap_mode("view")
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```
```{r}
tmap_mode("plot")
```
### Visualizing coefficient estimates

```{r}
tmap_options(check.and.fix = TRUE)
tmap_mode("view")
AREA_SQM_SE <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)
```
### By URA Planning Region

```{r}
tm_shape(mpsz[mpsz$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(gwr_sf_adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```

