---
title: "In-class Exercise 08"
author: "Ng Meng Ye"
---

# Geographically Weighted Predictive Models

Predictive modelling uses statistical learning or machine learning techniques to predict outcomes.

Random forest method.

### The data

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse,
               knitr, kableExtra)
```

Read the input data sets. It is in **simple feature data frame**.

```{r}
mdata <- read_rds("C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex08/data/mdata.rds")
```


Calibrating predictive models are computational intensive, especially random forest method is used. For quick prototyping, a 10% sample will be selected at random from the data by using the code chunk below.

```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(1500)
```

The code chunk below is used to check if there are overlapping point features.

```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

::: callout-note
When using GWmodel to calibrate explanatory or predictive models, it is very important to ensure that there are no overlapping point features
:::

In the code code chunk below, `st_jitter()` of sf package is used to move the point features by 5m to avoid overlapping point features.

```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 5)
```

results = round up if amount is too small all geometry will change, no just overlapped point you can check the difference before and after.

6 digits postal code, one building can have multiple points.

GWR cannot use same data point.

Split into training and test data, 65% for training, 35% for test `initial_split()` of **rsample**

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.67/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

Multicollinearity check
```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```

Building a non-spatial multiple linear regression

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```


Multicollinearity check with VIF

```{r}
vif <- performance::check_collinearity(price_mlr)
kable(vif, 
      caption = "Variance Inflation Factor (VIF) Results") %>%
  kable_styling(font_size = 18) 
```

```{r}
plot(vif) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Predictive Modelling with gwr

Computing adaptive bandwidth

```{r}
library(GWmodel)

```


```{r}
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```
```{r}
bw_adaptive
```



Model Calibration

```{r}
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```

Predictive Modelling with MLR
Test data BW
```{r}
gwr_bw_test_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```
```{r}
gwr_bw_test_adaptive
```


Predicting

```{r}
install.packages("GWmodel")
```


```{r}
library(GWmodel)
```



```{r}
gwr_pred <- gwr.predict(formula = resale_price ~
                          floor_area_sqm + storey_order +
                          remaining_lease_mths + PROX_CBD + 
                          PROX_ELDERLYCARE + PROX_HAWKER + 
                          PROX_MRT + PROX_PARK + PROX_MALL + 
                          PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                          WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
                          WITHIN_1KM_PRISCH, 
                        data=train_data, 
                        predictdata = test_data, 
                        bw=26, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)
```


Predictive Modelling: RF method

Data preparation

Firstly, code chunk below is used to extract the coordinates of training and test data sets
```{r}
coords <- st_coordinates(HDB_sample)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```
```{r}
train_data_nogeom <- train_data %>%
  st_drop_geometry()
```

Predictive Modelling: SpatialML method
```{r}
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + 
                       storey_order + remaining_lease_mths + 
                       PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + 
                       PROX_MRT + PROX_PARK + PROX_MALL + 
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data_nogeom, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

Predicting by using the test data

Preparing the test data
```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```

Predicting with the test data

In the code chunk below, predict.grf() of spatialML for predicting re-sale prices in the test data set (i.e. test_data_nogeom)

```{r}
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

Creating DF
Next, the code chunk below is used to convert the output from predict.grf() into a data.frame.
```{r}
GRF_pred_df <- as.data.frame(gwRF_pred)
```

Then, cbind() is used to append fields in GRF_pred_df data.frame onto test_data.
```{r}
test_data_pred <- cbind(test_data, 
                        GRF_pred_df)
```

Visualizing the predicted values
```{r}
ggplot(data = test_data_pred,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
