---
title: "Take-home Exercise 3a: Modelling Geography of Financial Inclusion with Geographically Weighted Methods"
author: "Ng Meng Ye"
date: 2024-11-09
---

## 1 Overview

Financial inclusion is the availability and equality of opportunities to access financial services.

## 2 Getting Started

### Objectives

In view of this, we need to build an explanatory model to determine factors affecting financial inclusion by using geographical weighted regression (GWR) methods.

### The Study Area

The focus of this study would in the Uganda, at the **district** level.

## 3 Data Preparation

Setting the analytical tools.

```{r}
#| output: false
pacman::p_load(sp, sf, st, spdep, raster, spatstat, tmap, devtools,vtable, ggplot2, egg, corrplot, ggstats, ggstatsplot, GWmodel, tidyverse, gtsummary, olsrr, car, blorr,ISLR, klaR, rsample,kableExtra, performance,readxl,dplyr)
```

### Data Preparation

-   geoBoundaries-UGA-ADM2, available publicly from [geoBoundaries](https://www.geoboundaries.org/countryDownloads.html), is a geospatial dataset in SHP format that delineates the district-level boundaries of our study area, Uganda.

-   FinScope Uganda 2023, available publicly from [FinScope](https://fsduganda.or.ug/data-sets-finscope-uganda-2023-survey-report/), is in XLSX format and includes survey data on financial inclusion among the adult population (aged 16 and above) in Uganda.

## 4 Data Wrangling

### Uganda Geospatial Data

Importing Geospatial Data into R

We will import **geoBoundaries-UGA-ADM3** `shp` file into RStudio and save the file into an R dataframe called `uganda`

```{r}
uganda <-  st_read(dsn = "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/",
                 layer = "geoBoundaries-UGA-ADM3")
```

Transform CRS to WGS 84 / UTM zone 36N (EPSG: 32636)

```{r}
uganda_32636 <- st_transform(uganda, crs = 32636)
```

Check the projected CRS is WGS 84 / UTM zone 36N

```{r}
st_geometry(uganda_32636)
```

Use *st_crs()* to display the coordinate system of `uganda_32636`

```{r}
st_crs(uganda_32636)
```

Plot the Uganda map

```{r}
plot(st_geometry(uganda_32636))
```

Based on the plot, there are islands. Need to make sure that the point is not located on the small islands.

Convert multipolygon to individual polygon

Calculate the area of each polygon

```{r}
sf_polygon <- uganda_32636 %>% 
  st_cast("POLYGON") %>% 
  mutate(area = st_area(.))
```

Group by the unique name and select the largest polygon by area

```{r}
uganda_modified <- sf_polygon %>% 
  group_by(shapeName) %>% 
  filter(area == max(area)) 
```

view the resulting data frame

```{r}
uganda_modified
```

Plot the uganda map

```{r}
plot(st_geometry(uganda_modified))
```

From the plot, we can see that the Lake Victoria is included, this should be removed.

View the list of district name

```{r}
unique(uganda_modified$shapeName)
```

Remove 'Lake Victoria' from the data frame.

```{r}
uganda_modified <- uganda_modified %>%
  filter(shapeName != "Lake Victoria")
```

Verify that 'Lake Victoria' is removed.

```{r}
unique(uganda_modified$shapeName)
```

Plot the map again

```{r}
plot(st_geometry(uganda_modified))
```

### FinScope Attribute Data

Importing the attribute data from **Finscope-2023_Dataset_Final.xlsx**

```{r}
finscope <- read_excel("C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/Finscope-2023_Dataset_Final.xlsx",sheet = "Final_Dataset")
```

Display the data structure

```{r}
glimpse(finscope)
```

See the data in XCOORD column

```{r}
head(finscope$hh_gps_longitude) #see the data in XCOORD column
```

See the data in YCOORD column

```{r}
head(finscope$hh_gps_latitude) #see the data in XCOORD column
```

Remove rows with empty longitude or latitude

```{r}

finscope <- finscope %>%
  filter(hh_gps_longitude != 0, hh_gps_latitude != 0)

```

Convert aspatial data frame into a sf object

```{r}
finscope.sf <- st_as_sf(finscope,
                            coords = c("hh_gps_longitude", "hh_gps_latitude"),
                            crs=4326) %>%
  st_transform(crs=32636)
```

List the content of `finscope.sf`

```{r}
head(finscope.sf)
```

Show the distinct value of district

```{r}
unique(finscope.sf$District)
```

Change to title case

```{r}
library(stringr)
finscope.sf$District <- str_to_title(finscope$District)
```

View the changes

```{r}
unique(finscope.sf$District)
```

Extract the district list from `finscope`

```{r}
district_list <- finscope.sf %>%
  distinct(District) %>%               
  mutate(District = str_to_title(District)) %>%  
  arrange(District)                    

```

Extract the district list from `uganda_modified`

```{r}
district_list2 <- uganda_modified %>%
  distinct(shapeName) %>%               
  arrange(shapeName) 
```

Combine both list together

```{r}
merged_list <- district_list2 %>%
  left_join(district_list %>% mutate(In_List = TRUE), by = c("shapeName" = "District")) %>%
  mutate(In_List = ifelse(is.na(In_List), FALSE, TRUE))
```

View the district with no data collected in `finscope`

```{r}
false_districts <- merged_list %>%
  filter(In_List == FALSE)

print(false_districts$shapeName)
```

Hightlight the district in the map

```{r}
highlight_districts <- merged_list %>%
  filter(In_List == FALSE) %>%
  pull(shapeName)

highlighted_map <- uganda_modified %>%
  filter(shapeName %in% highlight_districts)

plot(st_geometry(uganda_modified), col = "lightgray", main = "Districts with no data in Finscope in Uganda", border = "black")

plot(st_geometry(highlighted_map), col = "red", add = TRUE, border = "black")

text(st_coordinates(st_centroid(st_geometry(highlighted_map))), 
     labels = highlighted_map$shapeName, 
     cex = 0.7, pos = 3, col = "black")

```

## 5 Exploratory Data Analysis (EDA)

Categorize Age, one of the independent variables

```{r}
finscope.sf <- finscope.sf %>%
  mutate(age_group = cut(
    age,
    breaks = c(15, 19, 29, 39, 59, 79, 100),  # Start with 15 to include age 16
    labels = c("Teenagers", "Young Adults", "Adults", "Middle-aged Adults", "Seniors", "Elderly"),
    right = TRUE,  # Includes the right boundary in each interval
    include.lowest = TRUE  # Ensures the lowest value (16) is included
  ))


```

Plot the distribution of dependent variables and independent variables

```{r}
library(ggpubr)

plot_categorical <- function(data, column, title) {
  ggplot(data, aes(x = !!sym(column))) +
    geom_bar(fill = "skyblue") +
    labs(title = title, x = column, y = "Count") +
    theme_minimal()
}

plot_payment_wallets <- plot_categorical(finscope.sf, "h2_2_8", "Mobile Money / E-money Wallets Usage")
plot_insurance <- plot_categorical(finscope.sf, "j1", "Existing Insurance Policy")
plot_payment_bank <- plot_categorical(finscope.sf, "h2_2_5", "Mobile / Internet Banking Usage")

plot_education <- plot_categorical(finscope.sf, "c4", "Education Level")
plot_mobile <- plot_categorical(finscope.sf, "c7_1_1", "Own a Mobile Phone?")
plot_internet <- plot_categorical(finscope.sf, "c6_1_2", "Internet Access")
plot_sex <- plot_categorical(finscope.sf, "c2", "Sex")
plot_marital <- plot_categorical(finscope.sf, "c3", "Marital Status")
plot_age <- plot_categorical(finscope.sf, "age_group", "Age_group")

```

Dependent variables

```{r}
# Arrange and display dependent variable plots
dependent_plots <- ggarrange(
  plot_payment_wallets, plot_insurance, plot_payment_bank,
  ncol = 1, nrow = 3)
dependent_plots  # Display the dependent variable plots

```

Independent variable

```{r}
# Arrange and display independent variable plots
independent_plots <- ggarrange(
  plot_education, plot_mobile, plot_internet, plot_age, plot_sex, plot_marital,
  ncol = 2, nrow = 3)
independent_plots 
```

## 6 Geographically Weighted Logistic Regression

### Dependent variable: Mobile Money / E-money Wallets Usage

Remove empty values in `h2_2_8` field which represents the Mobile Money / E-money Wallets Usage

```{r}
finscope_mobile_money <- finscope.sf %>%
  filter(!is.na(h2_2_8) & h2_2_8 != "")
```

Recode **Mobile Money / E-money Wallets Usage** to 1 and 0

```{r}
finscope_mobile_money$h2_2_8 <- ifelse(finscope_mobile_money$h2_2_8 == 1, 1, 0)
```

Independent variable = Education level, Own a mobile phone?, internet access, age_group, sex, and marital status

Convert Categorical Variables to Factors

```{r}
# Convert categorical variables to factors if they are not already
finscope_mobile_money$c2 <- as.factor(finscope_mobile_money$age_group)     
finscope_mobile_money$c2 <- as.factor(finscope_mobile_money$c2)           # sex
finscope_mobile_money$c3 <- as.factor(finscope_mobile_money$c3)           # marital status
finscope_mobile_money$c4 <- as.factor(finscope_mobile_money$c4)           # education level
finscope_mobile_money$c7_1_1 <- as.factor(finscope_mobile_money$c7_1_1)   # own a mobile phone?
finscope_mobile_money$c6_1_2 <- as.factor(finscope_mobile_money$c6_1_2)   # internet access

```

```{r}
# Fit the logistic regression model
model <- glm(h2_2_8 ~ age_group + c2 + c3 + c4 + c6_1_2 + c7_1_1, 
             family = "binomial", data = finscope_mobile_money)

```

Explanation of the Model

```{r}
summary(model)
```

::: callout-note
The intercept represents the log-odds of the outcome (using mobile money/e-money wallets) when all predictors are at their reference levels.

**age_group** The reference level for age_group is likely Teenagers (since it’s not listed), and the other age groups are compared to this reference level.

age_group **Young Adults** (Estimate = 0.60199, p ≈ 0.068): Young adults have higher log-odds (0.60199) of using mobile money/e-money wallets compared to teenagers. Odds Ratio: exp(0.60199) ≈ 1.825, suggesting that young adults have approximately 1.83 times higher odds of using mobile money/e-money wallets than teenagers.

age_group **Adults**, age_group **Middle-aged Adults**, age_group **Seniors**: None of these age groups have statistically significant coefficients (p \> 0.05), meaning there’s no strong evidence of a difference in usage between these age groups and teenagers.

age_group **Elderly** (Estimate = -1.33603, p = 0.273): Elderly individuals have lower log-odds of using mobile money/e-money wallets compared to teenagers, but this result is not statistically significant.

**Sex (c2)** c22 **Male vs. Female**: The coefficient for males (0.11156) is not statistically significant (p = 0.454), suggesting that there is no strong evidence of a difference in usage between males and females.

**Marital Status** c32, c33: None of the levels of marital status are statistically significant, indicating no clear association between marital status and the likelihood of using mobile money/e-money wallets.

**Education Level** Reference Group: The baseline group is "Never went to school". All other education levels are compared to this group. Some primary completed (not P7):

Slightly higher odds of using mobile money/e-money wallets compared to those who never went to school. Odds Ratio: 1.55 (about 55% higher odds), but the result is only marginally significant. Completed P7, Some secondary completed (not S6), Completed S6, Specialized training/certificate, Specialized training or diploma, and Completed degree and above:

No significant difference in mobile money/e-money wallet usage compared to those who never went to school. "Don't know" (education level not specified):

Also shows no significant difference in usage compared to those who never went to school.

**Internet Access**

c6_1_22 (Internet Access = Yes): The coefficient is -0.11348 and is not statistically significant (p = 0.550), indicating no strong evidence that internet access influences the likelihood of using mobile money/e-money wallets in this dataset.

**Own a mobile phone** c7_1_12 (Own a Mobile Phone = Yes, Estimate = -1.49183, p \< 0.001): This coefficient is statistically significant and negative, suggesting that individuals who do not own a mobile phone have much lower odds of using mobile money/e-money wallets compared to those who do own a mobile phone. Odds Ratio: exp(-1.49183) ≈ 0.225, meaning that not owning a mobile phone is associated with approximately 0.23 times the odds (or a 77% reduction in odds) of using mobile money/e-money wallets, compared to those who own a mobile phone.
:::

#### Data Preparation for GWLR Model

Convert `finscope.sf` to spatial data

```{r}
finscope.sp <- as_Spatial(finscope_mobile_money)
```

Calculating distance matrix to fit the GWLR model

```{r}
distMAT <- gw.dist(dp.locat = coordinates(finscope.sp))
```

```{r}
dim(distMAT)
```

Computing Adaptive Bandwidth with dependent variable = Mobile Money / E-money Wallets Usage Independent variable = Sex

```{r}
bw.adaptive <- bw.ggwr(formula = h2_2_8 ~ c2, 
                       family = "binomial", 
                       data = finscope_mobile_money, 
                       approach = "CV", 
                       kernel = "gaussian", 
                       adaptive = TRUE, 
                       longlat = FALSE, 
                       p = 2, 
                       theta = 0, 
                       dMat = distMAT)

```

::: callout-note
The resulting adaptive bandwdith value is 626, represent the number of nearest neighbours.
:::

#### Fitting GWLR model

Set `bw` parameter to 626. The `family` parameter is set to "binomial" due to the binary nature of the `Mobile Money/E-money wallets usage` variable.

```{r}
gwlr <- ggwr.basic(h2_2_8 ~ age_group + c7_1_1 + c6_1_2 + c3,
                   data = finscope_mobile_money, 
                   bw = 626, 
                   family = "binomial", 
                   kernel = "gaussian", 
                   adaptive = TRUE, 
                   cv = T, 
                   tol = 1e-05, 
                   maxiter = 20, 
                   p = 2, 
                   theta = 0, 
                   longlat = FALSE, 
                   dMat = distMAT)
```

```{r}
gwlr
```

::: callout-note
Pseudo R-square (0.087): This value is relatively low, suggesting that the model explains around 8.7% of the variation in **Mobile Money / E-money Wallets Usage**

The decrease in AIC from the global model's AIC(1289.2) suggests that the GWLR model provides a slightly better fit due to spatial variation.

c7_1_1 (Owns a Mobile Phone - Yes): Strongly significant and negative (-1.480, p \< 0.001). Not owning a mobile phone is associated with much lower odds of using mobile money/e-money wallets.

c6_1_2 (Internet Access): Not significant (p = 0.99963), indicating no strong association between internet access and mobile money/e-money usage.

Education Level (c3): Only some levels of c3 (e.g., c34) are significant, indicating that education level might affect usage, though the impact varies across levels.
:::

#### Visualize GWR Output

converting SDF into sf data.frame

```{r}
mobilemoney.sf.adaptive <- st_as_sf(gwlr$SDF) %>% 
  st_transform(crs=32636)
```

```{r}
mobilemoney.sf.adaptive.UTM36N <- st_transform(mobilemoney.sf.adaptive, 32636)
mobilemoney.sf.adaptive.UTM36N  
```

```{r}
gwlr.output <- as.data.frame(gwlr$SDF)
mobilemoney.sf.adaptive <- cbind(finscope_mobile_money, as.matrix(gwlr.output))
```

Calculate local pseudo R² for each observation.

```{r}
# Step 1: Calculate the null deviance from a null model (intercept-only model)
null_model <- glm(h2_2_8 ~ 1, data = finscope_mobile_money, family = "binomial")
null_deviance <- deviance(null_model)

# Step 2: Extract local fitted values and residuals, converting to numeric if necessary
local_fitted_values <- as.numeric(mobilemoney.sf.adaptive$yhat)
local_residuals <- as.numeric(mobilemoney.sf.adaptive$residual)

# Step 3: Calculate Local Pseudo R² for each observation
mobilemoney.sf.adaptive$local_pseudo_r2 <- 1 - (local_residuals^2 / null_deviance)
```

```{r}
tmap_mode("view")
tm_shape(uganda_modified)+ 
  tm_polygons(alpha = 0.1) +
tm_shape(mobilemoney.sf.adaptive) +  
  tm_dots(col = "local_pseudo_r2", 
          border.col = "gray60",
          border.lwd = 1)
```

::: callout-note
In logistic regression (and similarly in geographically weighted logistic regression), a pseudo R² is an approximation of the proportion of variance in the dependent variable (in this case, mobile money/e-money wallet usage) explained by the model.

Higher Pseudo R² Values indicate areas where the model fits well.

Lower Pseudo R² Values (Lighter Colors)indicate areas where the model explains less of the variability in mobile money usage
:::

### Dependent variable: Mobile / Internet Banking Usage

Remove empty values in `h2_2_5` field which represents the Mobile Money / E-money Wallets Usage

```{r}
finscope_internet_banking <- finscope.sf %>%
  filter(!is.na(h2_2_5) & h2_2_5 != "")
```

Recode **Mobile / Internet Banking Usage** to 1 and 0

```{r}
finscope_internet_banking$h2_2_5 <- ifelse(finscope_internet_banking$h2_2_5 == 1, 1, 0)
```

Independent variable = Education level, Own a mobile phone?, internet access, age_group, sex, and marital status

Convert Categorical Variables to Factors

```{r}

finscope_internet_banking$c2 <- as.factor(finscope_internet_banking$age_group)     
finscope_internet_banking$c2 <- as.factor(finscope_internet_banking$c2)           # sex
finscope_internet_banking$c3 <- as.factor(finscope_internet_banking$c3)           # marital status
finscope_internet_banking$c4 <- as.factor(finscope_internet_banking$c4)           # education level
finscope_internet_banking$c7_1_1 <- as.factor(finscope_internet_banking$c7_1_1)   # own a mobile phone?
finscope_internet_banking$c6_1_2 <- as.factor(finscope_internet_banking$c6_1_2)   # internet access

```

Fit the logistic regression model using `glm`

```{r}
model_2 <- glm(h2_2_5 ~ age_group + c2 + c3 + c4 + c6_1_2 + c7_1_1, 
             family = "binomial", data = finscope_internet_banking)
```

Explanation of the Model

```{r}
summary(model_2)
```

::: callout-note
The logistic regression model assesses the likelihood of using internet or mobile banking based on demographic and behavioral factors. Most variables, including age groups, are not significant, suggesting limited predictive power for these factors individually. However, education level (specifically, category c34) is borderline significant (p ≈ 0.054), hinting that higher education may slightly increase the likelihood of internet banking usage. Internet access (c6_1_22) and mobile phone ownership (c7_1_12) are significant predictors, with negative coefficients, indicating that lacking internet access or a mobile phone significantly reduces the likelihood of using internet or mobile banking. The model has a modest fit, with a pseudo R² around 8.9% (Null deviance vs. Residual deviance), and an AIC of 914.09, suggesting room for additional or alternative predictors. Five levels of c2 (sex) were omitted due to singularities, indicating redundancy or perfect multicollinearity.
:::

#### Data Preparation for GWLR Model

Convert `finscope_internet_banking` to spatial data

```{r}
finscope_internet_banking.sp <- as_Spatial(finscope_internet_banking)
```

Calculating distance matrix to fit the GWLR model

```{r}
distMAT_internet_banking <- gw.dist(dp.locat = coordinates(finscope_internet_banking.sp))
```

```{r}
dim(distMAT_internet_banking)
```

Computing Adaptive Bandwidth

```{r}
bw.adaptive_internet_banking <- bw.ggwr(formula = h2_2_5 ~ c3, 
                       family = "binomial", 
                       data = finscope_internet_banking, 
                       approach = "CV", 
                       kernel = "gaussian", 
                       adaptive = TRUE, 
                       longlat = FALSE, 
                       p = 2, 
                       theta = 0, 
                       dMat = distMAT_internet_banking)

```

::: callout-note
The resulting adaptive bandwdith value is 555, represent the number of nearest neighbours.
:::

#### Fitting GWLR model

Set `bw` parameter to 555. The `family` parameter is set to "binomial" due to the binary nature of the `Mobile / Internet Banking Usage` variable.

```{r}
gwlr_internet_banking <- ggwr.basic(h2_2_5 ~ c3 + c6_1_2 + c7_1_1,
                   data = finscope_internet_banking, 
                   bw = 555, 
                   family = "binomial", 
                   kernel = "gaussian", 
                   adaptive = TRUE, 
                   cv = T, 
                   tol = 1e-05, 
                   maxiter = 20, 
                   p = 2, 
                   theta = 0, 
                   longlat = FALSE, 
                   dMat = distMAT_internet_banking)
```

```{r}
gwlr_internet_banking
```

::: callout-note
Internet Access (c6_1_22): The local coefficients for internet access are consistently negative across locations, ranging from -0.975 to -0.930, indicating a uniformly strong negative effect on usage.

Mobile Phone Ownership (c7_1_12): The effect of mobile phone ownership also varies, ranging from -1.26 to -0.90, with stronger negative effects in some areas, suggesting that lack of mobile phone access has a more significant impact in certain regions.

Education Levels (c3 series): These variables show varying effects, but their median and quartile values suggest limited impact overall, as indicated by the global results.

AIC (916.88) and Pseudo R² (0.064) for the GWLR model indicate a slightly improved fit over the global model, capturing spatial heterogeneity but still explaining only a modest portion of the variance.
:::

#### Visualize GWR Output

```{r}
internetbanking.sf.adaptive <- st_as_sf(gwlr_internet_banking$SDF) %>% 
  st_transform(crs=32636)
```

```{r}
internetbanking.sf.adaptive.UTM36N <- st_transform(internetbanking.sf.adaptive, 32636)
internetbanking.sf.adaptive.UTM36N  
```

```{r}
gwlr_internet_banking.output <- as.data.frame(gwlr_internet_banking$SDF)
internetbanking.sf.adaptive <- cbind(finscope_internet_banking, as.matrix(gwlr_internet_banking.output))
```

Calculate local pseudo R² for each observation.

```{r}
null_model_ib <- glm(h2_2_5 ~ 1, data = finscope_internet_banking, family = "binomial")
null_deviance_ib <- deviance(null_model)

local_fitted_values_ib <- as.numeric(internetbanking.sf.adaptive$yhat)
local_residuals_ib <- as.numeric(internetbanking.sf.adaptive$residual)

internetbanking.sf.adaptive$local_pseudo_r2 <- 1 - (local_residuals_ib^2 / null_deviance_ib)
```

Plot the map

```{r}
tmap_mode("view")
tm_shape(uganda_modified)+ 
  tm_polygons(alpha = 0.1) +
tm_shape(internetbanking.sf.adaptive) +  
  tm_dots(col = "local_pseudo_r2", 
          border.col = "gray60",
          border.lwd = 1)
```

::: callout-note
Higher Pseudo R² Values indicate areas where the model fits well.

Lower Pseudo R² Values (Lighter Colors)indicate areas where the model explains less of the variability in mobile money usage

Kampala’s location in the central region likely influences the results, as urban settings tend to have better infrastructure, making predictors like mobile phone and internet access more uniformly available. This can affect the local pseudo R² values and how well the model fits for internet/mobile banking usage patterns.
:::

### Dependent variable: Existing Insurance Policy

Remove empty values in `j1` field which represents the Insurance Policy ownership

```{r}
finscope_insurance <- finscope.sf %>%
  filter(!is.na(j1) & j1 != "")
```

Recode **Existing Insurance Policy** to 1 and 0

```{r}
finscope_insurance$j1 <- ifelse(finscope_insurance$j1 == 1, 1, 0)
```

Independent variable = Education level, Own a mobile phone?, internet access, age_group, sex, and marital status

Convert Categorical Variables to Factors

```{r}

finscope_insurance$c2 <- as.factor(finscope_insurance$age_group)     
finscope_insurance$c2 <- as.factor(finscope_insurance$c2)           # sex
finscope_insurance$c3 <- as.factor(finscope_insurance$c3)           # marital status
finscope_insurance$c4 <- as.factor(finscope_insurance$c4)           # education level
finscope_insurance$c7_1_1 <- as.factor(finscope_insurance$c7_1_1)   # own a mobile phone?
finscope_insurance$c6_1_2 <- as.factor(finscope_insurance$c6_1_2)   # internet access

```

Fit the logistic regression model using `glm`

```{r}
model_3 <- glm(j1 ~ age_group + c2 + c3 + c4 + c6_1_2 + c7_1_1, 
             family = "binomial", data = finscope_insurance)
```

Explanation of the Model

```{r}
summary(model_3)
```

::: callout-note
The model suggests that lack of internet access is associated with lower likelihood of having insurance, which may reflect barriers in accessing information and services related to insurance. Other predictors, including age and mobile phone ownership, do not appear to have a statistically significant effect on insurance ownership in this dataset. The singularities in certain categories highlight potential multicollinearity issues, which could be further explored to refine the model.
:::

#### Data Preparation for GWLR Model

Convert `finscope_insurance` to spatial data

```{r}
finscope_insurance.sp <- as_Spatial(finscope_insurance)
```

Calculating distance matrix to fit the GWLR model

```{r}
distMAT_insurance <- gw.dist(dp.locat = coordinates(finscope_insurance.sp))
```

```{r}
dim(distMAT_insurance)
```

Computing Adaptive Bandwidth

```{r}
bw.adaptive_insurance <- bw.ggwr(formula = j1 ~ c6_1_2, 
                       family = "binomial", 
                       data = finscope_insurance, 
                       approach = "CV", 
                       kernel = "gaussian", 
                       adaptive = TRUE, 
                       longlat = FALSE, 
                       p = 2, 
                       theta = 0, 
                       dMat = distMAT_insurance)

```

::: callout-note
The resulting adaptive bandwdith value is 387, represent the number of nearest neighbours.
:::

#### Fitting GWLR model

Set `bw` parameter to 387. The `family` parameter is set to "binomial" due to the binary nature of the `Existing Insurance Policy` variable.

```{r}
gwlr_insurance <- ggwr.basic(j1 ~ c3 + c6_1_2 + c7_1_1 + c2 + c3 + c4,
                   data = finscope_insurance, 
                   bw = 387, 
                   family = "binomial", 
                   kernel = "gaussian", 
                   adaptive = TRUE, 
                   cv = T, 
                   tol = 1e-05, 
                   maxiter = 20, 
                   p = 2, 
                   theta = 0, 
                   longlat = FALSE, 
                   dMat = distMAT_insurance)
```

```{r}
gwlr_insurance
```
::: callout-note
The GWR model’s pseudo R² is approximately 45%, a notable improvement over the global model’s fit, suggesting that the GWR model captures significant spatial variation in the predictors' effects.

AIC for GWR is 602.46, indicating a slightly higher AIC than the global model, likely due to the additional complexity introduced by modeling spatial variability.

Internet Access and Mobile Phone Ownership appear to have strong, location-dependent impacts on insurance ownership, indicating that these factors are crucial for facilitating access to insurance in certain regions.
Policy Implications: Regions with strong negative impacts from lack of internet or mobile phone access might benefit from policies aimed at improving digital infrastructure or providing digital literacy programs to enhance awareness and accessibility of insurance products.

Variable Influence by Region: The variability in coefficients suggests that the predictors’ influence on insurance adoption may depend on local characteristics, such as urbanization, socioeconomic factors, or access to insurance providers.
:::

#### Visualize GWR Output
```{r}
insurance.sf.adaptive <- st_as_sf(gwlr_insurance$SDF) %>% 
  st_transform(crs=32636)
```

```{r}
insurance.sf.adaptive.UTM36N <- st_transform(insurance.sf.adaptive, 32636)
insurance.sf.adaptive.UTM36N  
```


```{r}
gwlr_insurance.output <- as.data.frame(gwlr_insurance$SDF)
insurance.sf.adaptive <- cbind(finscope_insurance, as.matrix(gwlr_insurance.output))
```

Calculate local pseudo R² for each observation.

```{r}
null_model_in <- glm(j1 ~ 1, data = finscope_insurance, family = "binomial")
null_deviance_in <- deviance(null_model)

local_fitted_values_in <- as.numeric(insurance.sf.adaptive$yhat)
local_residuals_in <- as.numeric(insurance.sf.adaptive$residual)

insurance.sf.adaptive$local_pseudo_r2 <- 1 - (local_residuals_in^2 / null_deviance_in)
```

Plot the map
```{r}
tmap_mode("view")
tm_shape(uganda_modified)+ 
  tm_polygons(alpha = 0.1) +
tm_shape(insurance.sf.adaptive) +  
  tm_dots(col = "local_pseudo_r2", 
          border.col = "gray60",
          border.lwd = 1)
```
::: callout-note
The prevalence of darker colors suggests a strong and consistent fit of the GWLR model across Uganda, indicating that the selected predictors are likely important and relevant factors for insurance ownership nationwide. This offers confidence in the model's results and provides useful insights for policies targeting these factors across different regions.
:::

```{r}
write_rds(uganda_32636, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/uganda_32636.rds")
write_rds(uganda_modified, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/uganda_modified.rds")
write_rds(finscope_insurance, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/finscope_insurance.rds")
write_rds(finscope_internet_banking, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/finscope_internet_banking.rds")
write_rds(finscope_mobile_money, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/finscope_mobile_money.rds")
write_rds(model, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/model.rds")
write_rds(model_2, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/model_2.rds")
write_rds(model_3, "C:/ngmengye/ISSS626-GAA/Take-home_Ex/Take-home_Ex03/Uganda/data/rds/model_3.rds")
```
