---
title: "Hands-on Exercise 07"
author: "Ng Meng Ye"
---

# Geographically Weighted Regression

It fits a local regression model at each location rather than a global model for the entire dataset, enabling the model to capture spatial variations in the relationships between variables.

### The data

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

URA Master Plan 2014’s planning subzone boundaries

```{r}
mpsz = st_read(dsn = "C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex07/data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz_svy21)
```

```{r}
st_bbox(mpsz_svy21) #view extent
```

```{r}
condo_resale = read_csv("C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex07/data/aspatial/Condo_resale_2015.csv")
```

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

North side of equatorial

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

```{r}
summary(condo_resale)

```

converts `condo_resale` data frame into simple feature data frame by using `st_as_sf()` `st_transform()` to convert the coordinates from wgs84(crs=4326) to svy21(crs-3414)

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

`head()` is used to list the content of cond_resale.sf object

```{r}
head(condo_resale.sf)
```

## Exploratory Data Analysis (EDA)

Use histogram to show the distribution of SELLING_PRICE

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Normalize using log transformation

Derive a new variable called LOG_SELLING_PRICE

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

```{r}
ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Plot small multiple histograms, also known as **trellis plot**

`ggarrange()` is used to organize these histogram

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, 
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

Draw statistical point map

Review the geospatial distribution condominium resale prices in Singapore

```{r}
tmap_mode("view") #turn on the interactive mode of tmap
```
```{r}
tm_shape(mpsz_svy21)+tmap_options(check.and.fix = TRUE)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot") #turn R display into plot mode
```

##Hedonic Pricing Modelling in R

### Simple Linear Regression

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf) #SELLING_PRICE = dependent variable, AREA SQM = independent variable
```

`lm` returns an object of class 'lm' for multiple responses of class c("mlm","lm")

ANOVA test

```{r}
summary(condo.slr) #Print the Analysis of variance table of the results
```

Selling price = -258121.1 if the predicted value of SELLING_PRICE when the AREA_SQM is 0

For each additional square meter, the selling price increases by an estimated 14,719 SGD

p-value = 2e-16, indicating that the area has a highly significant effect on the selling price.

Residual standard error 942700 is in dollars, the measure units of the dependent variable. SGD942,700 on average, the actuall selling prices differ from the model's predictions.

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

### Multiple Linear Regression

Avoid multicollinearity

```{r}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE", #parameter order = AOE, angular order of the eigenvectors
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

**Freehold** is highly correlated to **LEASE_99YEAR**. Exclude **LEASE_99YEAR**

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

Remove those variables which are not statistically significant

1.  PROX_TOP_PRIMARY_SCH
2.  PROX_SUPERMARKET
3.  PROX_HAWKER_MARKET
4.  PROX_KINDERGARTEN

```{r}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                 data=condo_resale.sf)
ols_regress(condo.mlr1)
```
Prepare publication quality table: `gtsummary` method

```{r}
tbl_regression(condo.mlr1, intercept = TRUE) #tbl_regression() is used to create a well formatted regression report.
```

How to add model statistics?

use `add_glance_table()` to append to the report table
use `add_glance_source_note()` to add as a table source

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

Use `olsrr` to perform OLS regression

`ols_vif_tol()` to test multicollinearity

```{r}
ols_vif_tol(condo.mlr1)
```
Less than 10 = no sign of multicollinearity

`ols_plot_resid_fit()` to perform linearity assumption test

```{r}
ols_plot_resid_fit(condo.mlr1)
```
Data points scattered around line 0 = dependent and independent variables are linear

`ols_plot_resid_hist()` to perform normality assumption test

```{r}
ols_plot_resid_hist(condo.mlr1)
```
Residual of the multiple linear regression model = normal distribution


`ols_test_normality()` to perform formal statistical test methods
```{r}
ols_test_normality(condo.mlr1)
```
Visual the residual of the hedonic pricing model

Convert **condo_resale.sf** from sf data frame into a **SpatialPointsDataFrame**

Export residual of the hedonic pricing model and save it as a data frame.

```{r}
mlr.output <- as.data.frame(condo.mlr1$residuals)
```

Join the newly created data frame with **condo_resale.sf**

```{r}
condo_resale.res.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
```

Convert **condo_resale.res.sf** from simple feature object into a **SpatialPointsDataFrame**

```{r}
condo_resale.sp <- as_Spatial(condo_resale.res.sf)
condo_resale.sp
```
Display the distribution of the residuals

```{r}
tmap_mode("view")
```
```{r}
tm_shape(mpsz_svy21)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale.res.sf) +  
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```
The map reveal that there is a sign of spatial autocorrelation.

Perform Moran's I test to proof.

```{r}
nb <- dnearneigh(coordinates(condo_resale.sp), 0, 1500, longlat = FALSE)
summary(nb)
```
```{r}
nb_lw <- nb2listw(nb, style = 'W')
summary(nb_lw)
```

Perform Moran'I test for residual spatial autocorrelation.
```{r}
lm.morantest(condo.mlr1, nb_lw)
```
p-value < 0.05, reject the null hypothesis

### Use GWmodel to build Hedonic Pricing Models

Fixed Bandwidth

```{r}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                     PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                     PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                     PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                     FAMILY_FRIENDLY + FREEHOLD, 
                   data=condo_resale.sp, 
                   approach="CV", 
                   kernel="gaussian", 
                   adaptive=FALSE, 
                   longlat=FALSE)
```
```{r}
gwr.fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                         PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                         PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                         PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                         FAMILY_FRIENDLY + FREEHOLD, 
                       data=condo_resale.sp, 
                       bw=bw.fixed, 
                       kernel = 'gaussian', 
                       longlat = FALSE)
```


```{r}
gwr.fixed
```
Build adaptive bandwidth GWR model

```{r}
bw.adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale.sp, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```
```{r}
gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale.sp, bw=bw.adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE, 
                          longlat = FALSE)
```
```{r}
gwr.adaptive
```
convert SDF into sf data.frame

SDF = an object can contains fit.points, GWR coefficient estimates, y value, predicted values, coefficient standard errors and t-values

```{r}
condo_resale.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

```{r}
condo_resale.sf.adaptive.svy21 <- st_transform(condo_resale.sf.adaptive, 3414)
condo_resale.sf.adaptive.svy21  
```
```{r}
gwr.adaptive.output <- as.data.frame(gwr.adaptive$SDF)
condo_resale.sf.adaptive <- cbind(condo_resale.res.sf, as.matrix(gwr.adaptive.output))
```

```{r}
glimpse(condo_resale.sf.adaptive)
```
```{r}
summary(gwr.adaptive$SDF$yhat)
```
Visualizing local R2

```{r}
tmap_mode("view")
tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

Visualize coefficient estimates

```{r}
tmap_mode("view")
AREA_SQM_SE <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)
```


URA Planning Region
```{r}
tm_shape(mpsz_svy21[mpsz_svy21$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(condo_resale.sf.adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```

