---
title: "In-class Exercise 1"
author: "Ng Meng Ye"
date: "Aug 20 2024"
date-modified: "last-modified"
execute: 
  eval: true
echo: true
message: false
freeze: true
---

```{r}
library(sf)
pacman::p_load(sf,ggstatsplot)
```


mpsz14_kml = st_read("C:/ngmengye/ISSS626-GAA/In-class_exercise/In-class_ex01/Data/MasterPlan2014SubzoneBoundaryWebKML.kml")


The source file is corrupted

```{r}
preschool = st_read("C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex01/data/geospatial/PreSchoolsLocation.kml")
```


mpsz = st_read(dsn = "C:/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex01/data/geospatial", layer = "MP14_SUBZONE_WEB_PL")

mpsz3414 <- st_set_crs(mpsz, 3414) 
st_write(mpsz14_shp, ":/ngmengye/ISSS626-GAA/Hands-on_exercise/Hands-on_ex01/data/geospatial/MP14_SUBZONE_WEB_PL.kml", delete_dsn = TRUE)



```{r}
mpsz19_kml <- st_read("C:/ngmengye/ISSS626-GAA/In-class_exercise/In-class_ex01/Data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

```{r}
mpsz19_shp <- st_read(dsn = "C:/ngmengye/ISSS626-GAA/In-class_exercise/In-class_ex01/Data", layer = "MPSZ-2019")
```
```{r}
mpsz19_shp <- st_read(dsn = "C:/ngmengye/ISSS626-GAA/In-class_exercise/In-class_ex01/Data", layer = "MPSZ-2019")
                        
mpsz19_shp <- mpsz19_shp %>% st_transform(CRS=3414)

```
